---
title: "KHM Model"
author: "Mike"
date: "May 2, 2017"
output: html_document
  toc: toc1
---
## Overview

This is a KHM Project with the objective of predicting the final price of a home in Ames, Iowa based on 79 variables. 

To see further details on competition, go to:

https://www.kaggle.com/c/house-prices-advanced-regression-techniques#tutorials & 

Also, the following Kernal is also a useful resource to getting started:

https://www.kaggle.com/tannercarbonati/house-prices-advanced-regression-techniques/detailed-data-analysis-ensemble-modeling

## Load Datasets

```{r setup, include=FALSE, message=FALSE , warning= FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("C:/Users/AblMi001/Desktop/Kaggle Housing Market")
require(ggplot2) # for data visualization
require(stringr) #extracting string patterns
require(Matrix) # matrix transformations
require(glmnet) # ridge, lasso & elastinet
require(xgboost) # gbm
require(randomForest)
require(Metrics) # rmse
require(dplyr) # load this in last so plyr doens't overlap it
require(caret) # one hot encoding
require(scales) # plotting $$
require(e1071) # skewness
require(corrplot) # correlation plot
```

First step is to load and explore dataset. The competition is providing two different csv files:

Train: Training set  
Test: Test Set

```{r warning = FALSE, message= FALSE}

train <- read.csv("train.csv", stringsAsFactors =  FALSE)
test <- read.csv("test.csv", stringsAsFactors = FALSE)

```

Next, let's combine the dataset to expand our data size. For us to perform analysis at a more productive rate we can combine the 2 dataframes together and run analysis on both data sets at once, then split the data once we are ready to train a model.  

However, we'll need to remove the Id field and remove the SalesPrice in the train dataset. The `Id` feature is useless so we can toss it out of our dataset and we won't include `SalePrice` since it is our response variable. We also won't import string variables as factors since our ultimate goal is to tranform all our variables to numeric.  

```{r warning= FALSE, message= FALSE}
df.combined <- rbind(within(train, rm('Id', 'SalePrice')), within(test, rm('Id')))
dim(df.combined) 
```

## Missing Values

Our dataset is filled with many missing values, therefore, before we can build any predictive model we'll clean our data by filling in all NA's with appropriate values. For each column we'll try to replace NA's by using features that have a strong correlation, which will help us determine what values to fill in.  


```{r warning= FALSE, message= FALSE}
#identify which columns have missing values
na.cols <- which(colSums(is.na(df.combined)) > 0)
sort(colSums(sapply(df.combined[na.cols],is.na)), decreasing = TRUE)
paste('There are', length(na.cols), 'columns w/ mssing values')
```


```{r}
#helper function for plotting categoric data for easier data viz

plot.categoric = function(cols, df){
  for (col in cols) {
    order.cols = names(sort(table(df.combined[,col]), decreasing = TRUE))
  
    num.plot = qplot(df[,col]) +
      geom_bar(fill = 'cornflowerblue') +
      geom_text(aes(label = ..count..), stat='count', vjust=-0.5) +
      theme_minimal() +
      scale_y_continuous(limits = c(0,max(table(df[,col]))*1.1)) +
      scale_x_discrete(limits = order.cols) +
      xlab(col) +
      theme(axis.text.x = element_text(angle = 30, size=12))
  
    print(num.plot)
  }
}
```


**PoolQC: Pool Quality**
Of the 2909 houses with NA's for PoolQC only 3 recorded a PoolArea greater than 0. To fill the missing values for these houses we can see what the mean pool area is for each quality of pool. To do this we'll group by each unique vlaue of `PoolQC` and compute the mean of PoolArea for that quality. Once we have the mean areas computed we can assign the quality that has the closest corresponding mean of PoolArea

```{r warning= FALSE, message= FALSE}
df.combined[, c('PoolQC', 'PoolArea')] %>%
  group_by(PoolQC) %>%
  summarise(mean = mean(PoolArea), counts = n())
```

Note that only 10 houses total have pools  The rest of the houses PoolQC can be filled with 'NoPool' since each has a corresponding PoolArea of 0. We'll fill in the 3 PoolQC's from above with the quality having the closest corresponding mean. 

```{r warning= FALSE, message= TRUE}
df.combined[2421,'PoolQC'] = 'Ex'
df.combined[2504,'PoolQC'] = 'Ex'
df.combined[2600,'PoolQC'] = 'Fa'
df.combined$PoolQC[is.na(df.combined$PoolQC)] = 'None'
```

**GarageType: Garage location**  
**GarageYrBlt: Year garage was built**  
**GarageFinish: Interior finish of the garage**  
**GarageCars: Size of garage in car capacity**  
**GarageArea: Size of garage in square feet**  
**GarageQual: Garage quality**  
**GarageCond: Garage condition**  

Lets see what the deal is with `GarageYrBlt`. It seems reasonable that most houses would build a garage when the house itself was built. We can check this by seeing how many houses were built the same year their garage was built. 


```{r}
length(which(df.combined$GarageYrBlt == df.combined$YearBuilt))
```

2216 of the 2919 houses have same year for for GarageYrBlt and YearBuilt. Lets replace any of the NA's for `GarageYrBlt` with the year from  `YearBuilt`. 

```{r}
idx = which(is.na(df.combined$GarageYrBlt))
df.combined[idx, 'GarageYrBlt'] = df.combined[idx, 'YearBuilt']
```



That leaves 6 garage features in our dataset and 4 of them have at least 157 missing values while `GarageArea` and `GarageCars` both only have 1, thus we can assume this particular house does not have a garage at all. For the rest of the houses we can check to see that if the NA's recorded also have 0 GarageArea and 0 GarageCars. If they do we can fill in their missing values with 'None' since having 0 area and 0 cars in their garage will imply that they do not have any at all. 

```{r}
garage.cols = c('GarageArea', 'GarageCars', 'GarageQual', 'GarageFinish', 'GarageCond', 'GarageType')
df.combined[is.na(df.combined$GarageCond),garage.cols]
```

Only one house who had NA's in their garage columns had an area graeteer than 0. We can fill this house in manually and set the rest of the houses NA's to 0. 

For the house with *GarageArea = 360* and *GarageCars = 1*, but NA's in the other columns, we can use the most frequent values for each columns from houses with a similar area and car count. 

```{r warning=FALSE, message=FALSE}
idx = which(((df.combined$GarageArea < 370) & (df.combined$GarageArea > 350)) & (df.combined$GarageCars == 1))

names(sapply(df.combined[idx, garage.cols], function(x) sort(table(x), decreasing=TRUE)[1]))
  
df.combined[2127,'GarageQual'] = 'TA'
df.combined[2127, 'GarageFinish'] = 'Unf'
df.combined[2127, 'GarageCond'] = 'TA'
```


Now we can fill in any missing numeric values with 0 and categoric with 'None' since these houses recorded having 0 area and 0 cars in their garage.

```{r warning=FALSE, message=FALSE}
for (col in garage.cols){
  if (sapply(df.combined[col], is.numeric) == TRUE){
    df.combined[sapply(df.combined[col], is.na), col] = 0
  }
  else{
    df.combined[sapply(df.combined[col], is.na), col] = 'None'
  }
}
```


## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
